---
title: "Part 0: Setup"
author: Dan Turner (dturner@u.northwestern.edu)
---

# Part "0": **Setup instructions**

**Welcome to my tutorial on the basic of scraping the web using R!**

This is an *R Notebook* that will help you to make sure that your computer is ready for the workshop.

Run the code blocks by pressing the *Run* button while the curser is in that code chunk, or by the key combination *Cmd+Shift+Enter*. Each code chunk behaves kind of like an isolated `.R` script file, but the results appear beneath the code chunk, instead of in the console.

*Try it out with the code chunk below.*

```{r}

# What version of R are your running?
getRversion()

```

I have 3.6.2, but as long as you are above 3.2, you should be fine (otherwise, update now *https://www.r-project.org* and revisit these materials when you're done).

```{r RUN THIS}
library(rvest) # Web scraping
library(tidyverse) # Data wrangling
library(RCurl) # Downloading files from the internet
```

If there are no errors, then you are ready for the next part!

*What else do I need to start?*
You will also need to be ready to use a mainstream web browser like Firefox, Chrome, Safari, or Edge.

## Continue to Part 1: Scraping structured data
